### 2018-04-01 paper/sexSelectionSpacing-ver1.tex
Edit Results

I capture sex composition with dummy variables for the
possible combinations for the specific spell, ignoring the ordering of births.
As an example, for the third spell, three groups are used: Two boys,
one girl and one boy, and two girls.
As discussed, the sex composition of previous children affects both the timing
of births and the use of sex-selective abortions.

@@

First, did the spacing after girls increase relative to the spacing
after boys?
The underlying hypothesis is that in the absence of sex selection son 
preference leads to shorter spacing after the birth of a girl than after the 
birth of a boy, whereas son preference increases the spacing after the birth 
of a girl relative to after the birth of a boy when prenatal sex selection is 
available.
I still expect parents to \emph{try} to conceive earlier after the birth 
of a girl than after a boy, but the abortions will increase the 
length of the space, and may even lead to longer spacing after the
birth of girls than after boys.

Second, have the shapes of the survival curves changed?
The shapes of the survival curves provide clues both to general
changes in birth spacing and to whether parents are
still trying to conceive earlier after girls than boys.
I expect this to be the case if there is still a strong preference for
boys, and, as discussed above, this behavior may even become more pronounced
with access to sex selection.
At a technical level, the shapes of the survival curve serve as an
indicator for the presence of non-proportionality.

Finally, does the sex ratio change within a spell?
There are two reasons why this is interesting.
First, if the sex ratio changes within a spell, relying on the observed sex
ratio at birth---as is standard practice in the literature---will not provide
correct estimates of the use of sex selection or the sex ratio when 
childbearing is completed.
Second, any changes in the sex ratio within a spell provides clues
to parents' preferences over the ideal spacing between births and the
distribution of son preference in the sample.

% [How to read tables]
Tables \ref{tab:median_sex_ratio_low}, \ref{tab:median_sex_ratio_med}, and
\ref{tab:median_sex_ratio_high}
show predicted median duration and sex ratios by period, spell, and sex 
composition for the three different education levels separated by area of residence.%
\footnote{
Appendix Tables \ref{tab:p25_p75_low}, \ref{tab:p25_p75_med}, and 
\ref{tab:p25_p75_high} show 25th and 75th percentiles durations.
} 
The predictions are based on the estimation method described in 
Section \ref{sec:strategy}.
The standard errors for both measures are based on bootstrapping, where the 
original sample is resampled with replacement and the method is applied to 
the new sample.

The predicted median duration shown is the weighted average over the women in the 
sample's predicted median durations, using the individual probabilities of 
giving birth as weights.
For each woman in a sample, I calculate her probability of ending the spell 
with a birth, and her contribution to median spacing is how long it will take
from the beginning of the spell to where she is predicted to reach half of that probability.
For example, if a woman has an 80\% probability of having a birth by the end of 
the spell, her median duration is the number of months it takes to reach
40\% probability of having exited with a birth.
The 40\% is equivalent to the 60\% point in a standard survival curve because that 
captures how many have not yet exited.

For second through fourth spell I show whether durations for sex composition
other than only girls are statistically significantly different from the
duration with only girls.
I use bootstrapping to calculate the level of statistical significance.
For each new sample, I calculate the difference in duration and these
differences are used to calculate the standard error and the level of
significance for the difference in the original sample.

The cleanest test is comparing durations after only boys with durations after
only girls, but the number of births to women with only sons becomes small 
in the later periods because of falling fertility.
Hence, it is possible to have substantial differences in spacing that are
not statistically significant because of low power, especially for the third 
and fourth spell.
It is, in principle, possible to estimate the model taking into account the 
ordering of children, but this would further lower the power of the test, by 
adding one additional group for the third spell and four additional groups for the fourth spell.

The predicted sex ratio is the weighted average over the women in the sample's 
individual predicted sex ratios, using the individual probabilities of having
given birth by the end of the spell as weights.
Each woman's predicted sex ratio is the weighted average of the predicted
percentage boys over the months in the spell calculated using equation 
(\ref{eq:probability_son}) and the probability of giving birth in each month 
as weights.%
\footnote{
Imagine a spell has two periods and that the estimated percentage boys for a
woman are 54\% and 66\% and that the likelihood of having a birth is 20\% and 40\%.
The likelihood of having a birth is the change in the survival curve; 
in other words, there is 40\% chance that she will not have given birth by the end of 
this spell.
This woman's percentage boys is then $\frac{54*0.2+66*0.4}{0.2+0.4} = 62$.
}
The predicted sex ratio captures the percent boys that will be born to women 
in the sample once childbearing for that spell is over.

>>>>

I capture sex composition with dummy variables for the
possible combinations for the specific spell, ignoring the ordering of births.
As an example, for the third spell, three groups are used: Two boys,
one girl and one boy, and two girls.
It is, in principle, possible to estimate the model taking into account
the ordering of children, but this would further lower the power of the
test, by adding one additional group for the third spell and four
additional groups for the fourth spell. 
As discussed, the sex composition of previous children affects both the timing
of births and the use of sex-selective abortions.

@@

First, did the spacing after girls increase relative to the spacing
after boys?
The underlying hypothesis is that in the absence of sex selection son 
preference leads to shorter spacing after the birth of a girl than after the 
birth of a boy, whereas son preference increases the spacing after the birth 
of a girl relative to after the birth of a boy when prenatal sex selection is 
available.
I still expect parents to \emph{try} to conceive earlier after the birth 
of a girl than after a boy, but the abortions will increase the 
duration between births, and may even lead to longer spacing after the
birth of girls than after boys.

Second, have the shapes of the survival curves changed?
The shapes of the survival curves provide information both on general
changes in birth spacing and whether parents are
still trying to conceive earlier after girls than boys.
I expect this to be the case if there is still a strong preference for
boys, and, as discussed above, this behavior may even become more pronounced
with access to sex selection.

Finally, does the sex ratio change within a spell?
There are two reasons why this is interesting.
First, if the sex ratio changes within a spell, relying on the observed sex
ratio at birth---as is standard practice in the literature---will not provide
correct estimates of the use of sex selection or the sex ratio at
the conclusion of childbearing.
Second, any changes in the sex ratio within a spell provides clues
to parents' preferences over the ideal spacing between births and the
distribution of son preference in the sample.

% [How to read tables]
Tables \ref{tab:median_sex_ratio_low}, \ref{tab:median_sex_ratio_med}, and
\ref{tab:median_sex_ratio_high}
show the predicted median duration and sex ratios by period, spell, and sex 
composition for the three education levels separated by area of residence.%
\footnote{
Appendix Tables \ref{tab:p25_p75_low}, \ref{tab:p25_p75_med}, and 
\ref{tab:p25_p75_high} show 25th and 75th percentiles durations.
} 
For each subsample, I first estimate the model described in Section 
\ref{sec:strategy} and then use the estimated coefficients to predict 
median duration and sex ratio.
For the standard errors, I use bootstrapping for both measures,
where the model is repeatedly estimated using resampling with replacement.

The predicted median duration is the weighted average over the women in the 
subsample, using the individual probabilities of giving birth as weights.
For each woman in a sample, I calculate her probability of ending the spell 
with a birth, and her contribution to median spacing is how long it will take
from the beginning of the spell to where she is predicted to reach half of that probability.
For example, if a woman has an 80\% probability of having a birth by the end of 
the spell, her median duration is the number of months it takes to reach
40\% probability of having exited with a birth.
The 40\% is equivalent to the 60\% point on a standard survival curve because that 
captures how many have not yet exited.

For the second through the fourth spell, I show whether durations for sex 
composition other than only girls are statistically significantly different 
from the duration with only girls.
To calculate the level of statistical significance, I bootstrap the
difference in spacing.

The cleanest test is comparing durations after only boys with durations after
only girls, but the number of births to women with only sons becomes small 
in the later periods.
Hence, it is possible to have substantial differences in spacing that are
not statistically significant because of low power, especially for the third 
and fourth spell.

The predicted sex ratio is the weighted average over the women in the sample's 
individual predicted sex ratios, using the individual probabilities of having
given birth by the end of the spell as weights.
Each woman's predicted sex ratio is the weighted average of the predicted
percentage boys over the months in the spell calculated using equation 
(\ref{eq:probability_son}) and the probability of giving birth in each month 
as weights.%
\footnote{
Imagine a spell has two periods and that the estimated percentage boys for a
woman are 54\% and 66\% and that the likelihood of having a birth is 20\% and 40\%.
The likelihood of having a birth is the change in the survival curve; 
in other words, there is 40\% chance that she will not have given birth by the end of 
this spell.
This woman's percentage boys is then $\frac{54*0.2+66*0.4}{0.2+0.4} = 62$.
}
The predicted sex ratio captures the percent boys that will have been born to women 
in the sample when childbearing for that spell is over.
